---
title: "Gear Avoidance Behavior Simulation (gabs)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Some instructions:
# https://kbroman.org/pkg_primer/pages/vignettes.html
```

```{r setup}
library(gabs)
```

# Introduction

This vignette is a work in progress. Suggestions are welcome. Check back for updates.

In fisheries stock assessments, catch is thought of as a product of fishing effort 
and the catchability and availability of the species at the fishing site 
(Ricker 1975). Availability is defined as either the number or biomass of fish 
present at a site and catchability describes the rate at  which fish are caught 
relative to the true population size. In order to monitor changes in fish 
populations monitoring programs use catch per unit of effort (CPUE) as an index.
CPUE is proportional to the part of the population that is vulnerable to the 
type of gear that is used in sampling; the vulnerable part of the population 
is not necessarily the entire population (Maunder et al. 2006). This variable 
catchability creates a problem for interpreting monitoring data because CPUE 
values that are assumed to represent a fixed portion of the population represent 
the combination of catchability and availability. There is no direct way to 
separate these two effects in CPUE monitoring data without generating additional 
information.

There are methods to deal with varying catchability, such as calculating ratios 
and fitting statistical models that allow environmental conditions or time variables 
to account for changes in catchability (Maunder & Punt 2004). This package takes
a different approach. By simulating the behavior of individual fish in response
to approaching sampling gear, we can see the effects of varying catchability on
a known number of available fish.

A complete description of the formulas used in this package is documented in 
this paper: https://doi.org/10.20944/preprints202002.0177.v1

# Example
Put an example here!


# References

Maunder MN, Sibert JR, Fonteneau A, Hampton J, Kleiber P, & Harley SJ (2006) 
Interpreting catch per unit effort data to assess the status of individual stocks 
and communities. ICES Journal of Marine Sciences 63:1373-1385. 
DOI: 10.1016/j.icesjms.2006.05.008

Maunder MN & Punt AE (2004) Standardizing catch and effort data: a review of 
recent approaches. Fisheries Research 70:141-159. DOI: 10.1016/j.fishres.2004.08.002

Ricker WE (1975) Computation and interpretation of biological statistics of fish 
populations. Fisheries Research Board of Canada Bulletin 191.

